Optimal (Multiway) Spatial Joins

### 2024-05-25 msj.tex
yufei

Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$. Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle. We use $B_\bm{t}$ to represent $\bigcap_{i=1}^t r_i$, namely, the intersection of the rectangles in $\bm{t}$. Note that $B_\bm{t}$ is a rectangle itself. We define:
\myitems{
    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.

@@

\extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v) \in \J(R_1, ..., R_{k-3}, H', V)$. Note that $\bm{t}[k-2] = h'$ must intersect with $v$, because $\bm{t}$ is a result tuple of the $(k-1)$-SJ on $R_1, ..., R_{k-3}, H', V$. Suppose $h = [x_1, x_2] \times y$ and $p = (x, y)$ is the intersection point between $h$ and $v$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment $h' = [x_1, x] \times y$.
\myeqn{
    \contained_{R_{k-2}}(\bm{t}) &=& \{r \in R_{k-2} \mid \text{ the effective horizontal } \nn \\
    && \text{segment of $\bm{t}$ is contained in $r$}   \}
    \label{eqn:type2:contained-t}
}
The above should not be confused with \eqref{eqn:contained}, where the ``$\contained$'' function takes a segment as the input, rather than a tuple.

@@

We prove the next lemma in Appendix~\ref{app:hv}.

@@

\extraspacing {\bf Set $\bm{H^*}$.} 
For each segment $h' \in H'$, define
\myeqn{
    \minleft(h') = \min_{\substack{\text{$\bm{t} \in \J(R_1, ..., R_{k-3}, H', V):$}\\ \text{$\bm{t}[k-2] = h'$}}} \text{$x$-coordinate of $\bm{t}[k-1]$}. 
    \label{eqn:minleft}
}
We reminder the reader that, in the above equation, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment. We set $\minleft(h')$ to $\infty$, if no $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ has $h'$ in its field $\bm{t}[k-2]$. When $\minleft(h') \ne \infty$, assuming $h' = [x_1, x_2] \times y$, we introduce a horizontal segment
\myeqn{
    h^* &=& [x_1, \minleft(h')] \times y. \nn
}
and call it the {\em minimal segment} of $h'$.

@@

\extraspacing {\bf Computing the ``$\bm{\contained_{R_{k-2}}(t)}$'' Sets.} We are ready to produce $\contained_{R_{k-2}}(h^*)$, defined in \eqref{eqn:type2:contained-t}, for any $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, based on Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*}. First, fetch the (already computed) minimal segment $h^*$ of $\bm{t}[k-2]$ in $O(1)$ time. Then, scan the rectangles $r$ in $\contained_{R_{k-2}}(h^*)$ in descending order of $\xright(r)$. For each $r$ scanned, check whether $r \in 
\contained_{R_{k-2}}(\bm{t})$, or equivalently, whether $r$ covers the intersection point between $h$ and $v$; this takes $O(1)$ time. Abort the scan as soon as $r \notin 
\contained_{R_{k-2}}(\bm{t})$. This way, we compute $\contained_{R_{k-2}}(t)$ in $O(k + |\contained_{R_{k-2}}(t)|)$ time. Doing so for all $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ takes $O(k \cdot |\J| + \sum_\bm{t} |\contained_{R_{k-2}}(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.

>>>>

Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$. Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle. We define 
\myeqn{
    B_\bm{t} &=& \cap_{i=1}^t r_i \label{eqn:B_t}
}
namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Let us introduce:
\myitems{
    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.

@@

\extraspacing {\bf Generating $\bm{\J_2}$.} Take any $(k-1)$-tuple $\bm{t} = (r_1, ..., r_{k-3}, h', v) \in \J(R_1, ..., R_{k-3}, H', V)$. Observe that $B_\bm{t}$ --- defined in \eqref{eqn:B_t} --- is a point (due to $h'$ and $v$). Suppose $h' = [x_1, x_2] \times y$ and $B_\bm{t} = (x, y)$, we define the {\em effective horizontal segment} of $\bm{t}$ as the horizontal segment $[x_1, x] \times y$.

@@

Define
\myeqn{
    \contained_{R_{k-2}}(\bm{t}) &=& \{r \in R_{k-2} \mid \text{ the effective horizontal } \nn \\
    && \text{segment of $\bm{t}$ is contained in $r$}   \}
    \label{eqn:type2:contained-t}
}
The above should not be confused with \eqref{eqn:contained}, where the ``$\contained$'' function takes a segment as the input, rather than a tuple.

We prove the next lemma in Appendix~\ref{app:hv} (to avoid confusion, the reader may want to be reminded that, for each $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, $\bm{t}[k-2]$ is a horizontal segment and $\bm{t}[k-1]$ is a vertical segment).

@@

\extraspacing {\bf Set $\bm{H^*}$.} 
For each segment $h' \in H'$, define
\myeqn{
    \minleft(h') = \min_{\substack{\text{$\bm{t} \in \J(R_1, ..., R_{k-3}, H', V):$}\\ \text{$\bm{t}[k-2] = h'$}}} \text{$x$-coordinate of $\bm{t}[k-1]$}. 
    \label{eqn:minleft}
}
We set $\minleft(h')$ to $\infty$, if no $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$ has $h'$ in its field $\bm{t}[k-2]$. When $\minleft(h') \ne \infty$, assuming $h' = [x_1, x_2] \times y$, we introduce a horizontal segment
\myeqn{
    h^* &=& [x_1, \minleft(h')] \times y. \nn
}
and call it the {\em minimal segment} of $h'$.

@@

\extraspacing {\bf Computing the ``$\bm{\contained_{R_{k-2}}(t)}$'' Sets.} We are ready to produce $\contained_{R_{k-2}}(h^*)$, defined in \eqref{eqn:type2:contained-t}, for any $\bm{t} \in \J(R_1, ..., R_{k-3}, H', V)$, based on Statement (2) of Lemma~\ref{lmm:hv:type2:contained-h*}. First, fetch the (already computed) minimal segment $h^*$ of $\bm{t}[k-2]$ in $O(1)$ time. Then, scan the rectangles $r$ of $\contained_{R_{k-2}}(h^*)$ in descending order of $\xright(r)$. For each $r$ scanned, check whether $r \in \contained_{R_{k-2}}(\bm{t})$, or equivalently, whether $r$ covers $B_\bm{t}$ (recall that $B_\bm{t}$ is a point); the cost for such a check is $O(1)$. Abort the scan as soon as $r \notin 
\contained_{R_{k-2}}(\bm{t})$. This way, $\contained_{R_{k-2}}(t)$ can be decided in $O(k + |\contained_{R_{k-2}}(t)|)$ time. Doing so for all $\bm{t} \in \J(R_1, ..., R_{k-3},$ $H', V)$ takes $O(k \cdot |\J| + \sum_\bm{t} |\contained_{R_{k-2}}(t)|) = O(k \cdot \out)$ time, where the derivation used Lemma~\ref{lmm:hv:type2:recur-output} and statement (3) of Lemma~\ref{lmm:hv:type2:properties}.
